AUTOMATED TRANSCRIPTION OF DRUMSOUNDS

### 2017-08-16 Latex/Paper/main.tex
Continued writing text about algorithm.

% Template for ICASSP-2016 paper; to be used with:
%          spconf.sty  - ICASSP/ICIP LaTeX style file, and
%          IEEEbib.bst - IEEE bibliography style file.
% --------------------------------------------------------------------------
\documentclass{article}
\usepackage{spconf,amsmath,graphicx}
\usepackage{blindtext}

@@

In the field of music information retrieval NMF has also been used in a variety of applications.
Most notable are the fields of source separation and music transcription.
For source separation the goal is to separate a piece of music containing a mix of multiple sound sources into the their components.
Virtanen \cite{virtanen2007} has researched this topic among many others.

@@

As mentioned before we used a non-negative matrix factorization to extract the time points of drum events from a drum track.
Generally speaking the NMF factorizes a matrix (usually called $\mathbf{V}$) into two smaller matrices (usually called $\mathbf{W}$ and $\mathbf{H}$) with the constraint that every matrix element must be non-negative.
In turn the sum of $\mathbf{W}$ and $\mathbf{H}$ is an approximation of $\mathbf{V}$ (see figure \ref{fig:NMF}).

@@

\caption{Matrix $\mathbf{V}$ can be factorized into two matrices $\mathbf{W}$ and $\mathbf{H}$, so that their sum approximates $\mathbf{V}$. \scriptsize{\textcopyright WikiMedia CC BY-SA 3.0}}
\label{fig:NMF}

@@

For our use case we were not interested in finding all the components and their activations.
We were rather interested in finding the corresponding activation of a specific component, like that of a base drum or a hi-hat.
This is also possible using NMF.
In order to do so, we specified $\mathbf{W}$ as the spectrum of a base drum and $\mathbf{V}$ continued to be the spectrum of the whole song.
The NMF then solves for the matrix $\mathbf{H}$ which tells at what point in time the specified base drum spectrum appears in the song.
In that setup the NMF essentially works as a pattern matching algorithm.

@@

\caption{(a) Magnitude spectrum of a drum track. (b) Magnitude spectrum of a base drum. (c) Activation matrix of the base drum. (d) Approximated spectrum. \scriptsize{\textcopyright Own work CC BY-SA 3.0}}
\label{fig:MatrixOverview}

@@

>>>>

% Template for ICASSP-2016 paper; to be used with:
%          spconf.sty  - ICASSP/ICIP LaTeX style file, and
%          IEEEbib.bst - IEEE bibliography style file.
% --------------------------------------------------------------------------
\documentclass{article}
\usepackage{spconf,amsmath,graphicx}
\usepackage{blindtext}
\usepackage[utf8]{inputenc}

@@

In the field of music information retrieval NMF has also been used in a variety of applications.
Most notable are the fields of source separation and music transcription.
For source separation the goal is to separate a piece of music containing a mix of multiple sound sources into the their components.
T. Virtanen \cite{virtanen2007} among many others has researched this topic.

In the field of music transcription NMF has also been used for different topics.
Smaragdis and Brown \cite{smaragdis2003} used NMF to analyze and extract polyphonic, harmonic content from a piece of music.
In the sub-field of drum transcription there has also been some research in the recent years.
Paulus and Virtanen \cite{paulus2005} as well as Moreau and Flexer \cite{moreau2007} as well as Wu and Lerch \cite{wu2015} have all shown that NMF is a technique that can be well applied to the problem of finding the time points of percussive events in music.

@@

As mentioned above we used a non-negative matrix factorization to extract the time points of drum events from a drum track.
In general terms the NMF factorizes a matrix (usually called $\mathbf{V}$) into two smaller matrices (usually called $\mathbf{W}$ and $\mathbf{H}$) with the constraint that every matrix element must be non-negative.
In turn the sum of $\mathbf{W}$ and $\mathbf{H}$ is an approximation of $\mathbf{V}$ (see figure \ref{fig:NMF}).

@@

\caption{Matrix $\mathbf{V}$ can be factorized into two matrices $\mathbf{W}$ and $\mathbf{H}$, so that their sum approximates $\mathbf{V}$. \scriptsize{\textsf{\textcopyright} WikiMedia CC BY-SA 3.0}}
\label{fig:NMF}

@@

For our use case we were not interested in finding all the components and their activations.
We were rather interested in finding the corresponding activation of a specific component, like that of a base drum or a hi-hat.
This is also possible using NMF.
In order to do so, we specified $\mathbf{W}$ as the spectrum of a base drum (see figure \ref{fig:MatrixOverview} (b)) and $\mathbf{V}$ continued to be the spectrum of the whole song (see figure \ref{fig:MatrixOverview} (a)).
The NMF then solves for the matrix $\mathbf{H}$ (see figure \ref{fig:MatrixOverview} (c)) which tells us at what point in time the specified base drum spectrum appears in the song.
In that setup the NMF essentially works as a pattern matching algorithm.

@@

\caption{(a) Magnitude spectrum of a drum track. (b) Magnitude spectrum of a base drum. (c) Activation matrix of the base drum. (d) Approximated spectrum. \scriptsize{\textsf{\textcopyright} Own work CC BY-SA 3.0}}
\label{fig:MatrixOverview}

\begin{minipage}[b]{1.0\linewidth}
  \centering
  \centerline{\includegraphics[width=8.5cm]{figures/Flowchart}}
%  \vspace{2.0cm}
  \medskip
\end{minipage}

\caption{Flowchart of the whole drum transcription algorithm. \scriptsize{\textsf{\textcopyright} Own work CC BY-SA 3.0}}
\label{fig:Flowchart}

Figure \ref{fig:Flowchart} shows a flowchart with the complete overview of the algorithm we developed.
On the left hand side the two inputs to the algorithms are shown.
The first input is a time domain signal of percussive instrument, like a base drum or hi-hat, and the second input is a time domain signal of a piece of music.
For both inputs the first step is to apply the short-time Fourier transform (STFT) which transforms both signal from the time domain to the frequency domain.
For the parameters of the STFT we used a frame size of 1024 samples which yields in a frequency resolution of approx. 43 Hz per bin.
We also chose an overlap of 50\% between hops to improve the time resolution.
As mentioned before a visual representations of both transformed inputs can be found in figure \ref{fig:MatrixOverview} (b) and (a) respectively.
Before both inputs together are fed into the actual detection the drum sound undergoes some further processing.
This step is marked as template extraction in figure \ref{fig:Flowchart}.
%% TODO: explain transient and shit
After both inputs are prepared they are handed over to the NMF, which does its processing in the way described in the paragraph above.
As explained the result of the NMF is in our case the activation matrix $\mathbf{H}$.
The matrix $\mathbf{H}$ can also be viewed as a function curve of how much the given template is contribution to the given music signal over time.
An example of the representation of the matrix $\mathbf{H}$ as a curve can be seen in figure \ref{fig:ActivationMatrix}.

\begin{minipage}[b]{1.0\linewidth}
  \centering
  \centerline{\includegraphics[width=8.5cm]{figures/ActivationMatrix}}
%  \vspace{2.0cm}
  \medskip
\end{minipage}

\caption{Example of the representation of an activation matrix as a curve. \scriptsize{\textsf{\textcopyright} Own work CC BY-SA 3.0}}
\label{fig:ActivationMatrix}

The next step of the algorithm is called onset detection where the exact time points are extracted from that curve.
The onset detection is a simple implementation loosely based on the algorithms described by Lerch \cite{lerch2012book} in chapter 6.3.
To extract the onsets of the drum events the significant local maxima of the curve (peaks) have to be detected.
In order to do that the curve is first normalized.
Afterwards a threshold of $\frac{2}{3}$ is applied.
This value was establish empirically to separate the valid detection from false positives.
Then the first derivative, which represents the slope of the curve, is calculated, by subtracting each element from the previous one.
An example of the resulting curve can be found in figure \ref{fig:Ableitung}.

\begin{minipage}[b]{1.0\linewidth}
  \centering
  \centerline{\includegraphics[width=8.5cm]{figures/Ableitung}}
%  \vspace{2.0cm}
  \medskip
\end{minipage}

\caption{An example of the first derivative of the activation function. \scriptsize{\textsf{\textcopyright} Own work CC BY-SA 3.0}}
\label{fig:Ableitung}

@@

Improvements: 
- Using adaptive templates like shown in \cite{lerch2015} can help the "over-fitting" of a template.
- adaptive threshold instead of 2/3
